---
title: "Bootstrap"
output: rmdformats::robobook
date: "2022-12-6"
---

```{r}
library('ggplot2')
```

```{r}
df.state.history <- read.csv('data/RDC_Inventory_Core_Metrics_State_History.csv')
df.DC <- df.state.history[df.state.history$state_id=='DC',]
```

# Check outliers using boxplots

```{r}
#png('Boxplot of Average Listing Price.png',bg='transparent')
boxplot(df.DC$average_listing_price, ylab = "Average listing price")
#dev.off()

#png('Boxplot of Median Listing Price per square foot.png',bg='transparent')
boxplot(df.DC$median_listing_price_per_square_foot, ylab = "Median Listing Price per square foot")
#dev.off()
```

# Delete outlier

```{r}
max.avg.price <- max(df.DC$average_listing_price)
print(max.avg.price)
df.DC <- df.DC[df.DC$average_listing_price!=max.avg.price,]
```
```{r}
df.DC$year = substr(df.DC$month_date_yyyymm,1,4)
```

# Check distribution of average_listing_price before and after pandemic

```{r}
#png('Distribution of Average Listing Price (2016-2019).png',bg='transparent')
hist(df.DC$average_listing_price[which(df.DC$year<=2019)],prob=T,main='Distribution of Average Listing Price (2016-2019)',xlab='Average Listing Price')
#dev.off()

#png('Distribution of Average Listing Price (2020-2022).png',bg='transparent')
hist(df.DC$average_listing_price[which(df.DC$year>2019)],prob=T,main='Distribution of Average Listing Price (2020-2022)',xlab='Average Listing Price')
#dev.off()
```

# Check distribution of Median Listing Price/sqft before and after pandemic

```{r}
#png('Distribution of Median Listing Price per square foot (2016-2019).png',bg='transparent')
hist(df.DC$median_listing_price_per_square_foot[which(df.DC$year<=2019)],prob=T,main='Distribution of Median Listing Price per square foot (2016-2019)',xlab='Median Listing Price (per square foot)')
#dev.off()

#png('Distribution of Median Listing Price per square foot (2020-2022).png',bg='transparent')
hist(df.DC$median_listing_price_per_square_foot[which(df.DC$year>2019)],prob=T,main='Distribution of Median Listing Price per square foot (2020-2022)',xlab='Median Listing Price (per square foot)')
#dev.off()
```


# Normality tests of average_listing_price before and after pandemic

```{r}
shapiro.test(df.DC$average_listing_price[which(df.DC$year<=2019)])
shapiro.test(df.DC$average_listing_price[which(df.DC$year>2019)])
```

# Normality tests of median_listing_price_per_square_foot before and after pandemic

```{r}
shapiro.test(df.DC$median_listing_price_per_square_foot[which(df.DC$year<=2019)])
shapiro.test(df.DC$median_listing_price_per_square_foot[which(df.DC$year>2019)])
```

# Bootstrap sample mean difference

```{r}
# average_listing_price
N <- 10000
diff.sample <- rep(NA,N)

for (i in 1:N){
  boot.sample.mean.before <- mean(sample(df.DC$average_listing_price[which(df.DC$year<=2019)], length(df.DC$average_listing_price[which(df.DC$year<=2019)]), replace=T))
  
  boot.sample.mean.after <- mean(sample(df.DC$average_listing_price[which(df.DC$year>2019)], length(df.DC$average_listing_price[which(df.DC$year>2019)]), replace=T))
  
  diff.sample[i] <- boot.sample.mean.after - boot.sample.mean.before
}

(diff.mean <- mean(diff.sample))

(q.95 <- quantile(diff.sample, c(.025, .975)))
```

```{r}
# median_listing_price_per_square_foot
N <- 10000
diff.sample1 <- rep(NA,N)

for (i in 1:N){
  boot.sample.mean.before1 <- mean(sample(df.DC$median_listing_price_per_square_foot[which(df.DC$year<=2019)], length(df.DC$median_listing_price_per_square_foot[which(df.DC$year<=2019)]), replace=T))
  
  boot.sample.mean.after1 <- mean(sample(df.DC$median_listing_price_per_square_foot[which(df.DC$year>2019)], length(df.DC$median_listing_price_per_square_foot[which(df.DC$year>2019)]), replace=T))
  
  diff.sample1[i] <- boot.sample.mean.after1 - boot.sample.mean.before1
}

(diff.mean1 <- mean(diff.sample1))

(q.951 <- quantile(diff.sample1, c(.025, .975)))
```

# Distribution of sample mean diffences

```{r}
#png('Distribution of sample mean differences for Average Listing Price.png',bg='transparent')
hist(diff.sample,prob=T,main='Distribution of sample mean diffences for Average Listing Price',xlab='Diffences')
abline(v=c(q.95[1],diff.mean,q.95[2]),col=c('blue','green','red'))
#dev.off()
```

```{r}
#png('Distribution of sample mean diffences for median unit price.png',bg='transparent')
hist(diff.sample1,prob=T,main='Distribution of sample mean diffences for median unit price',xlab='Diffences')
abline(v=c(q.951[1],diff.mean1,q.951[2]),col=c('blue','green','red'))
#dev.off()
```

