---
title: "data_testing"
author: "Victor De Lima"
date: "2022-11-29"
output: html_document
---



Data Science question: 
What are the variables contributing to the prediction of median days on market in the properties in the DC, Virginia and Maryland areas?




Bring in the data

```{r}

inventory_core_metrics_raw <-read.csv("./data/RDC_Inventory_Core_Metrics_Zip.csv")

head(inventory_core_metrics_raw)

```

Cleaning

### Inventory Core Metrics


```{r}
# Remove last row containing dataset description
inventory_core_metrics_clean <- head(inventory_core_metrics_raw, - 1)
```

```{r}
#Verify only a single time period is present
print("Time periods present:")
length(unique(inventory_core_metrics_clean$month_date_yyyymm ))
table(inventory_core_metrics_clean$month_date_yyyymm)

#Verify all post codes appear just once
print("Post Codes All Unique:")
length(unique(inventory_core_metrics_clean$postal_code)) == sum(table(inventory_core_metrics_clean$postal_code))

```

```{r}
# Remove month over month and year over year columns
inventory_core_metrics_clean = subset(inventory_core_metrics_clean, select = -c(
  median_listing_price_mm,
  #active_listing_count_yy,
  price_increased_count_mm,
  #price_reduced_count_yy,
  median_square_feet_mm,
  #average_listing_price_yy,
  #median_listing_price_yy,
  new_listing_count_mm,
  #price_increased_count_yy,
  median_listing_price_per_square_foot_mm,
  #median_square_feet_yy,
  pending_ratio_mm,
  median_days_on_market_mm,
  #new_listing_count_yy,
  pending_listing_count_mm,
  #median_listing_price_per_square_foot_yy,
  total_listing_count_mm,
  #pending_ratio_yy,
  active_listing_count_mm,
  #median_days_on_market_yy,
  price_reduced_count_mm,
  #pending_listing_count_yy,
  average_listing_price_mm
  #total_listing_count_yy
))

```



```{r}
#inventory_core_metrics_test <- inventory_core_metrics[inventory_core_metrics$postal_code %in% dc_zip_codes, ]

# regions
dc_metarea <- c(', dc', ', va',', md')

inventory_core_metrics_dcvamd <- inventory_core_metrics_clean[
  grep(paste(dc_metarea,collapse="|"), inventory_core_metrics_clean$zip_name), ]

inventory_core_metrics_dcvamd

```


```{r}

# Remove categorical columns
inventory_core_metrics = subset(inventory_core_metrics_dcvamd, select = -c(
  month_date_yyyymm,
  postal_code,
  zip_name
  
))

head(inventory_core_metrics)

```


EDA

```{r}
str(inventory_core_metrics)

```

```{r}
# Check amount of NAs
sapply(inventory_core_metrics, function(y) sum(length(which(is.na(y)))))

```


```{r}
# Drop columns with excessive missingness
inventory_core_metrics <- subset(inventory_core_metrics, select = -c(
  pending_listing_count,
  pending_ratio,
  price_increased_count_yy,
  price_reduced_count_yy,
  new_listing_count_yy,
  pending_listing_count_yy,
  pending_ratio_yy,
  median_days_on_market_yy, # this is the dependent variable
  median_listing_price_yy # showed multicollinearity

))

# Drop other missing values
#inventory_core_metrics <- inventory_core_metrics[!is.na(inventory_core_metrics$median_days_on_market),]
inventory_core_metrics <- na.omit(inventory_core_metrics)


```



```{r}
# Check amount of NAs
sapply(inventory_core_metrics, function(y) sum(length(which(is.na(y)))))

```



```{r}
dim(inventory_core_metrics)

```


Some variables may be insignificant in groups but significant when alone with the dependent variable. This happens when the correlation between X's is high. 


```{r}
library(ggcorrplot)

corr <- cor(inventory_core_metrics)
corr
ggcorrplot(corr)
```




```{r}
# Drop columns with excessive missingness
inventory_core_metrics <- subset(inventory_core_metrics, select = -c(
  total_listing_count,
  average_listing_price,
  price_reduced_count,
  new_listing_count

))

# Check correlations again

corr <- cor(inventory_core_metrics)
corr
ggcorrplot(corr)

```


```{r}
dim(inventory_core_metrics)

```



Create training and testing sets:

```{r}
# Reference: class material
library(caret)
set.seed(123)

training_samples <- createDataPartition(inventory_core_metrics$median_days_on_market, p = 0.8, list = FALSE)

inventory_core_metrics_train  <- inventory_core_metrics[training_samples, ]
inventory_core_metrics_test <- inventory_core_metrics[-training_samples, ]
dim(inventory_core_metrics_train)
dim(inventory_core_metrics_test)

```



Fitting the full regression:


```{r}
# Reference: class material
options(scipen=999)


regression_1 = lm(median_days_on_market~.,data=inventory_core_metrics_train)
regression_1_summary <- summary(regression_1)
regression_1_summary
```

Before going forward, let's check for multicollinearity to know if we have correlation between variables that can affect the regression fit. We use the Variance Inflation factor (VIF):

```{r}

library(car)

vif(regression_1)

```


```{r}

# Check interactions:
regression_1_interactions <- lm(median_days_on_market~(median_listing_price + active_listing_count + active_listing_count_yy  + price_increased_count + median_listing_price_per_square_foot + median_listing_price_per_square_foot_yy + median_square_feet + median_square_feet_yy + average_listing_price_yy + total_listing_count_yy + quality_flag)^2, data=inventory_core_metrics_train)
summary(regression_1_interactions)



```


Insignificant variables:

median_listing_price

```{r, message=FALSE, warning=FALSE}
# median_listing_price
library(ggplot2)
ggplot(data = inventory_core_metrics_train, aes(y = median_days_on_market, x = median_listing_price)) + 
  geom_point() + 
  labs(x = "Median Listing Price", y = "Median Days on Market", title = "Median Days on Market vs Median Listing Price") + 
  geom_smooth(method = 'lm') + 
  scale_y_continuous(limits = c(0, 200)) + 
  scale_x_continuous(limits = c(0, 1000000))

```

active_listing_count

```{r, message=FALSE, warning=FALSE}
# active_listing_count
ggplot(data = inventory_core_metrics_train, aes(y = median_days_on_market, x = active_listing_count)) + 
  geom_point() + 
  labs(x = "Active Listing Count", y = "Median Days on Market", title = "Median Days on Market vs Active Listing Count") + 
  geom_smooth(method = 'lm') + 
  scale_y_continuous(limits = c(0, 200)) + 
  scale_x_continuous(limits = c(0, 100))
```

```{r}
inventory_core_metrics_train_without_0_alc = inventory_core_metrics_train[inventory_core_metrics_train$active_listing_count != 0, ]
check_alc <- lm(median_days_on_market ~ log(active_listing_count), data = inventory_core_metrics_train_without_0_alc)
summary(check_alc)
```

```{r, message=FALSE, warning=FALSE}
# log(active_listing_count)
ggplot(data = inventory_core_metrics_train, aes(y = median_days_on_market, x = log(active_listing_count))) + 
  geom_point() + 
  labs(x = "Log of Active Listing Count", y = "Median Days on Market", title = "Median Days on Market vs Log of Active Listing Count") + 
  geom_smooth(method = 'lm')
```


median_days_on_market_yy


price_increased_count

```{r, message=FALSE, warning=FALSE}
# price_increased_count
ggplot(data = inventory_core_metrics_train, aes(y = median_days_on_market, x = price_increased_count)) + 
  geom_point() + 
  geom_smooth(method = 'lm') + 
  labs(x = "Price Increased Count", y = "Median Days on Market", title = "Median Days on Market vs Price Increased Count")
```
```{r}
check_pic <- lm(median_days_on_market ~ price_increased_count, data = inventory_core_metrics_train)
summary(check_pic)
```

median_listing_price_per_square_foot

```{r, message=FALSE, warning=FALSE}
# median_listing_price_per_square_foot
ggplot(data = inventory_core_metrics_train, aes(y = median_days_on_market, x = median_listing_price_per_square_foot)) + 
  geom_point() + 
  geom_smooth(method = 'lm') + 
  labs(x = "Median Listing Price per Square Foot", y = "Median Days on Market", title = "Median Days on Market vs Median Listing Price per Square Foot") +
  scale_y_continuous(limits = c(0, 300)) + 
  scale_x_continuous(limits = c(0, 700))
```

median_listing_price_per_square_foot_yy

```{r, message=FALSE, warning=FALSE}
# median_listing_price_per_square_foot_yy
ggplot(data = inventory_core_metrics_train, aes(y = median_days_on_market, x = median_listing_price_per_square_foot_yy)) + 
  geom_point() + 
  geom_smooth(method = 'lm') + 
  labs(x = "Median Listing Price per Square Foot yy", y = "Median Days on Market", title = "Median Days on Market vs Median Listing Price per Square Foot yy") +
  scale_y_continuous(limits = c(0, 200)) + 
  scale_x_continuous(limits = c(0, 2))
```

median_square_feet_yy

```{r, message=FALSE, warning=FALSE}
# median_square_feet_yy
ggplot(data = inventory_core_metrics_train, aes(y = median_days_on_market, x = median_square_feet_yy)) + 
  geom_point() + 
  geom_smooth(method = 'lm') + 
  labs(x = "Median Square Feet yy", y = "Median Days on Market", title = "Median Days on Market vs Median Square Feet yy") +
  scale_y_continuous(limits = c(0, 200)) + 
  scale_x_continuous(limits = c(0, 2))
```



```{r, message=FALSE, warning=FALSE}
# active_listing_count
ggplot(data = inventory_core_metrics_train, aes(y = median_days_on_market, x = active_listing_count_yy)) + 
  geom_point() + 
  labs(x = "Active Listing Count yy", y = "Median Days on Market", title = "Median Days on Market vs Active Listing Count yy") + 
  geom_smooth(method = 'lm') +
  scale_y_continuous(limits = c(0, 300)) + 
  scale_x_continuous(limits = c(0, 4))
```

median_square_feet

```{r, message=FALSE, warning=FALSE}
# median_square_feet
ggplot(data = inventory_core_metrics_train, aes(y = median_days_on_market, x = median_square_feet)) + 
  geom_point() + 
  geom_smooth(method = 'lm') + 
  labs(x = "Median Square Feet", y = "Median Days on Market", title = "Median Days on Market vs Median Square Feet") +
  scale_y_continuous(limits = c(0, 300)) + 
  scale_x_continuous(limits = c(0, 5000))
```

average_listing_price_yy

```{r, message=FALSE, warning=FALSE}
# average_listing_price_yy
ggplot(data = inventory_core_metrics_train, aes(y = median_days_on_market, x = average_listing_price_yy)) + 
  geom_point() + 
  geom_smooth(method = 'lm') + 
  labs(x = "Average Listing Price yy", y = "Median Days on Market", title = "Median Days on Market vs Average Listing Price yy") +
  scale_y_continuous(limits = c(0, 200)) + 
  scale_x_continuous(limits = c(0, 2))
```

total_listing_count_yy

```{r, message=FALSE, warning=FALSE}
# total_listing_count_yy
ggplot(data = inventory_core_metrics_train, aes(y = median_days_on_market, x = total_listing_count_yy)) + 
  geom_point() + 
  geom_smooth(method = 'lm') + 
  labs(x = "Total Listing Count yy", y = "Median Days on Market", title = "Median Days on Market vs Total Listing Count yy") +
  scale_y_continuous(limits = c(0, 300)) + 
  scale_x_continuous(limits = c(0, 2.5))
```





```{r}
inventory_core_metrics_train_without_0_alc = inventory_core_metrics_train[inventory_core_metrics_train$active_listing_count != 0, ]
check_alc <- lm(median_days_on_market ~ log(active_listing_count) + ., data = inventory_core_metrics_train_without_0_alc)
summary(check_alc)
```





```{r}
# Reference: class material
regression_data_1 <- subset(inventory_core_metrics_train, select = -c(median_listing_price, price_increased_count, median_listing_price_per_square_foot, median_listing_price_per_square_foot_yy, median_square_feet_yy, total_listing_count_yy, average_listing_price_yy))

regression_data_1 <- lm(median_days_on_market~.,data=regression_data_1)

regression_summary_1 <- summary(regression_data_1)
regression_summary_1
```





```{r}
# Reference: class material
regression_data_2 <- lm(median_days_on_market~

median_listing_price                                        +
active_listing_count                                        +
active_listing_count_yy                                     +
price_increased_count                                       +
median_listing_price_per_square_foot                        +
median_listing_price_per_square_foot_yy                     +
median_square_feet                                          +
median_square_feet_yy                                       +
average_listing_price_yy                                    +
total_listing_count_yy                                      +
quality_flag                                                +
median_listing_price*active_listing_count_yy                +
median_listing_price*median_listing_price_per_square_foot_yy+
median_listing_price*total_listing_count_yy                 +
active_listing_count_yy*median_listing_price_per_square_foot+
median_listing_price_per_square_foot*median_square_feet     +
median_listing_price_per_square_foot*total_listing_count_yy +
median_square_feet*median_square_feet_yy                    +
median_square_feet*average_listing_price_yy                 

, data=inventory_core_metrics_train)

regression_summary_2 <- summary(regression_data_2)
regression_summary_2

```

Let's also try having interaction terms and removing insignificant variables, as permissible by the hierarchy principle.

```{r}
# Reference: class material
regression_data_3 <- lm(median_days_on_market~

median_listing_price                                        +
active_listing_count                                        +
active_listing_count_yy                                     +
median_listing_price_per_square_foot                        +
median_listing_price_per_square_foot_yy                     +
median_square_feet                                          +
median_square_feet_yy                                       +
total_listing_count_yy                                      +
quality_flag                                                +
median_listing_price*active_listing_count_yy                +
median_listing_price*median_listing_price_per_square_foot_yy+
median_listing_price*total_listing_count_yy                 +
active_listing_count_yy*median_listing_price_per_square_foot+
median_listing_price_per_square_foot*median_square_feet     +
median_listing_price_per_square_foot*total_listing_count_yy +
median_square_feet*median_square_feet_yy 

, data=inventory_core_metrics_train)

regression_summary_3 <- summary(regression_data_3)
regression_summary_3

```


```{r}
# Reference: class material
regression_data_4 <- lm(median_days_on_market~

median_listing_price                                        +
log(active_listing_count)                                   +
active_listing_count_yy                                     +
median_listing_price_per_square_foot                        +
median_listing_price_per_square_foot_yy                     +
median_square_feet                                          +
median_square_feet_yy                                       +
total_listing_count_yy                                      +
quality_flag                                                +
median_listing_price*active_listing_count_yy                +
median_listing_price*median_listing_price_per_square_foot_yy+
median_listing_price*total_listing_count_yy                 +
active_listing_count_yy*median_listing_price_per_square_foot+
median_listing_price_per_square_foot*median_square_feet     +
median_listing_price_per_square_foot*total_listing_count_yy +
median_square_feet*median_square_feet_yy 

, data=inventory_core_metrics_train_without_0_alc)

regression_summary_4 <- summary(regression_data_4)
regression_summary_4

```


Getting $R^2$, $RSE$, $RMSE$ and $F-statistic$:

```{r, warning=FALSE, message=FALSE}
library(tidyverse)

# Original
predictions0 <- regression_1 %>% predict(inventory_core_metrics_test)
original <- c(
regression_1_summary$adj.r.squared,
regression_1_summary$sigma,
regression_1_summary$fstatistic[1],
RMSE(predictions0, inventory_core_metrics_test$danceability)
)

# Model 1
predictions1 <- regression_data_1 %>% predict(inventory_core_metrics_test)
model_1 <- c(
regression_summary_1$adj.r.squared,
regression_summary_1$sigma,
regression_summary_1$fstatistic[1],
RMSE(predictions1, inventory_core_metrics_test$median_days_on_market)
)

# Model 2
predictions2 <- regression_data_2 %>% predict(inventory_core_metrics_test)
model_2 <- c(
regression_summary_2$adj.r.squared,
regression_summary_2$sigma,
regression_summary_2$fstatistic[1],
RMSE(predictions2, inventory_core_metrics_test$median_days_on_market)
)

# Model 3
predictions3 <- regression_data_3 %>% predict(inventory_core_metrics_test)
model_3 <- c(
regression_summary_3$adj.r.squared,
regression_summary_3$sigma,
regression_summary_3$fstatistic[1],
RMSE(predictions3, inventory_core_metrics_test$median_days_on_market)
)

# Model 4
inventory_core_metrics_test_without_0_alc = inventory_core_metrics_test[inventory_core_metrics_test$active_listing_count != 0, ]
predictions4 <- regression_data_4 %>% predict(inventory_core_metrics_test_without_0_alc)
model_4 <- c(
regression_summary_4$adj.r.squared,
regression_summary_4$sigma,
regression_summary_4$fstatistic[1],
RMSE(predictions4, inventory_core_metrics_test_without_0_alc$median_days_on_market)
)

column_names <- c('Adj R Squared', 'RSE', 'F Statistic', 'RMSE')
eval_table <- rbind(original, model_1, model_2, model_3, model_4)
colnames(eval_table) <- column_names
eval_table

```


When comparing the original models, Model 3 is better in most used metrics. 

```{r}
par(mfrow=c(2,2))
plot(regression_data_3)
```

However, with the log term, we can obtain a slightly better result.

```{r}
par(mfrow=c(2,2))
plot(regression_data_4)
```




```{r}
anova(regression_data_1, regression_data_2)
```

```{r}
anova(regression_data_2, regression_data_3)
```


